---
title: ".large[Robust data analysis]"
subtitle: "An introduction to R"
author: "Sina Rüeger"
date: "2019-09-05"
output:
  xaringan::moon_reader:
    css: ["xaringan-themer.css", "extra-theme.css"]
    lib_dir: libs
    nature:
      slideNumberFormat: "%current%"                  # 12
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      highlightSpans: true
      titleSlideClass: [center, middle, inverse]
    self_contained: true
---



```{r xaringan-themer, include = FALSE}
library(xaringanthemer)
mono_light(
  base_color = "black",
  header_font_google = google_font("Arapey"), ## Oswald
  text_font_google   = google_font("Karla", "400", "700i"),  ## Roboto
  code_font_google   = google_font("Inconsolata")
  )
```


```{r, setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(knitr.kable.NA = '')
library(knitr)
library(kableExtra)
library(magrittr)
library(stringr)
library(ggplot2)

library(dplyr)
theme_set(hrbrthemes::theme_ipsum(subtitle_size = 14, caption_size = 9, axis_title_size = 12)) ## make sure to 

```

<!---------------------- SLIDE info ------------------------------>

<!--- adding a footer https://github.com/yihui/xaringan/wiki/Footer-and-header-lines ------>
---
class: left, middle

## 1. Download Slides: bit.ly/....

## 2. Go to Github repository: github.com/...

<!---------------------- ToC ------------------------------>

---
class: left, middle

.pull-left[
# 1. 
### What makes a Data Analysis robust?

<br>

# 3. 
### Workshop

]

.pull-right[
# 2. 
### How can R help with that?

<br>

# 4. 
### Learning R
]


<!------------------------------------------------------->
<!---------------------- PART 1 ------------------------->
<!------------------------------------------------------->
---
class: left, middle, inverse

# Part 1
# Data Analysis


---
class: left, middle

# Process to analyse data

- wickham+grolemund: circle (Wickham & Grolemund, 201)
- Elements & Principles (Hicks & Peng, 2019)

---
class: left, middle

# Loop
(wickham + grolemund)


---
class: left, middle

# Elements & Principles
(hicks + peng)

---
class: left, middle

# Overlap between both approaches

- tidy data
- reproducibility
- flow + exit the flow or repeat the flow, modify
- share results (which can then be used as an input again)

---
class: left, middle

# Requirements to a Software

- reproducible
- used by others
- continuously development
- high-level
- exchange results with others

<!------------------------------------------------------->
<!--------------- Thought experiment -------------------->
<!------------------------------------------------------->
---
class: left, middle, inverse

# Thought experiment


---
class: left, top

## Q1: 
### What is the most annoying & repetitive task at work? 
--
<br>
## Q2: 
### How could you automate it? 

---
class: left, middle

# Automation

Programming is essentially an automation of a process
- for repeatable tasks (like adding two numbers)
- scalable (many times repeating )
- complicated tasks (variable selection)
- where precision is needed (adding two decimals)
- reusing of frameworks (long code into functions)
- reproduce it

A side effect of a certain automation also brings reproducibility. 


<!------------------------------------------------------->
<!---------------------- PART 2 ------------------------->
<!------------------------------------------------------->
---
class: left, middle, inverse

# Part 2
# R

---
class: left, middle

# Capabilities of R

(show diagram from heidi seibold)
- It has been developed to do statistical data analysis
- Data insights
- simple calculations
- advanced statistical models
- statistical data analysis
- exchange with others
- [access API](https://medium.com/epfl-extension-school/an-illustrated-introduction-to-apis-10f8000313b9) (_So the API is a layer of code that sits between the database and most database users._)
- [produce an API](https://medium.com/tmobile-tech/r-can-api-c184951a24a3) (also [these resouces (h/t Sharla Gelfand)](https://twitter.com/sharlagelfand/status/1157794847438364679))
- read in data
- automise
- speed up
- run on a server
- talk to github
- enforces standards (because you need to work a certan way with a certain structure)

---
class: left, middle

### R has three things that I believe makes it special

- APIs
- RMarkdown (shiny, html, websites, ... )
- Plotting, anything visual

---
class: left, middle

# Constantly improving

There are things R is good at and tasks that R is not that good at. And this has changed over time. Because R is an open source project, the language is constantly evolving & improving. 


---
class: left, middle

# Functions and libraries

- Programming typically works like this: you apply a function to an object. 

- You could also say that you apply an action (a verb) to an object (a noun). 

- Excel also has functions: these are the tabs, and the object would be your active sheet. 

- This is an important propertiy, because a chain of such verbs can create a programming script. 

- If you hear the term library or package, then this is simply a collection of functions. In opensource languages, anyone can contribute a package. And we will do this later. 

- Example with plot: function, arguments, pacakage, libraary, objects


---
class: left, middle

.pull-left[
# R

- core
- anyone can contribute a package
- Show print screen / website

]


.pull-right[
# RStudio

- UI
- works for any platform
- costs for companies
- solutions for servers
- develop a lot of new packages
- anything between teaching and solutions for production
- shiny
- git : https://happygitwithr.com/
- Where are things in RStudio?: https://twitter.com/RLadiesNCL/status/1138812826917724160/photo/1

]

---
class: left, middle

# Analogy

R helps you with the grammar for your analsysis, but the content you are still responsible.


---
class: left, middle

# Hierarchy of best practices

1. documentation
1. run all script
1. version control
1. unit tests and sanity checks
1. write functions, package them and tell everyone
1. continuous integration
1. makefile (caching)

---
class: left, middle

# R in the wild

- https://www.hvitfeldt.me/blog/circle-love-making-hearts-with-circles/
- https://twitter.com/AmeliaMN/status/1126891572807729152?s=09
- sports science: https://jacquietran.github.io/2019_may_rladies_akl/R/#1
- ggplot2: https://cedricscherer.netlify.com/2019/05/17/the-evolution-of-a-ggplot-ep.-1/
- research compendium: 
    - https://github.com/seabbs/DirectEffBCGPolicyChange
    - https://github.com/venexia/MR-antihypertensives-AD
    - https://github.com/saralpulit/fatdistnGWAS
    - https://www.samabbott.co.uk/
- video, animation, presentation, website, write to excel




<!------------------------------------------------------->
<!---------------------- PART 3 ------------------------->
<!------------------------------------------------------->
---
class: left, middle, inverse

# Part 4
# Workshop

---
class: left, middle

Use paper + pen


---
class: left, middle


# 1. Plotting + model

## Task
Change something on the plot

## Extra
look into plotly, animation


---
class: left, middle


# 2. Make a map

## Task
Change pin location

## Extra
make a leaflet with watercolors



---
class: left, middle

# 3. Import data

Rexcel

## Task
import other data

## info
- datapasta: https://twitter.com/dataandme/status/1148548556850114561
- fabricatr: https://rviews.rstudio.com/2019/07/01/imagine-your-data-before-you-collect-it/


## extra
- different r packages to import data: readr, data.table
- create xls file


---
class: left, middle


# 4. Rmarkdown
export as html, combine 1 + 2 + 3
use janitor + skimr

## Task
export as pdf/word/markdown

## Infos
- https://rmd4sci.njtierney.com/
- rmarkdown: https://holtzy.github.io/Pimp-my-rmd/


## Extra

- Look into xaringan and CSS file
- code download button: https://twitter.com/apreshill/status/1108925218850893832


---
class: left, middle

# Commenting & Documentation

- How much documentation do you need? 
  - Think about your future self
  - Add simple instructions and ask a colleague to do it
- For structure use simple commands (align, lintr)
- A good exercise: look at an old script and reformat.
- [Reprohack!](https://sheffield-university.shinyapps.io/ReproHack_CCMcr/)





---
class: left, middle

# 5. Create an R package

create a package with corporate colors EPFL: https://www.epfl.ch/campus/services/communication/wp-content/uploads/2019/03/EPFL-brand-guidelines.pdf

- https://nanx.me/ggsci/
- color package: https://drsimonj.svbtle.com/creating-corporate-colour-palettes-for-ggplot2

## Task

Add an extra color 

## Info
- packages: https://www.slideshare.net/EmilyRobinson52/everything-you-wanted-to-know-about-making-an-r-package-but-were-afraid-to-ask
- packages: https://masalmon.eu/2017/12/11/goodrpackages/

## Extra 
Add your own function



---
class: left, middle

# 6. Extra stuff

- Shiny
- collaborative writing
- Git & R


---
class: left, middle

# Extra: Git & R

You use git already: congratulations. 

You want to use git but you are scared: 
- use the interface
- use gist.github.com (great for blogposts and small scripts)
- start with sharing slides / gist.github (from julie lowndes keynote)
- happygit with R
- https://saghirb.github.io/shared/Getting-Started-with-Git-and-GitHub-for-R-Users.html#1
- Playground: https://learngitbranching.js.org/




<!------------------------------------------------------->
<!---------------------- PART 4 ------------------------->
<!------------------------------------------------------->
---
class: left, middle, inverse


# Part 4
# Learning R


---
class: left, middle

# Strategies

- Learn isolated / digestable pieces
- surroud yourself with the language (fun examples + podcasts)
- use paper + pen
- tidytuesday, advent of code / tidies of march (irene steves)
- listen to podcast or watch videos: https://www.rstats.nyc/2019/nyr/

---
class: left, middle

# Embrace imperfection

Programming is 
- an iterative process
- and there is never only one solution

Do:
- Pair programming
- Code review
- Refactoring scripts of others


---
class: left, middle

# Asking others for help

- reprex
- SO, R-help, Rstuido help
- twitter


---
class: left, middle

# Become part of the community

- blogposts (rweekly!)
- R user group (RUG, R-Ladies)
- list: https://jumpingrivers.github.io/meetingsR/r-user-groups.html (if you know a meetup, make a PR)
- CoC
- R-Ladies
- French [slack](https://r-grrr.slack.com/join/shared_invite/enQtMzI4MzgwNTc4OTAxLWZlOGZiZTBiMWU0NDQ3OTYzOGE1YThiODgwZWNhNWEyYjI4ZDJiNmNhY2YyYWI5YzFiOTFkNDYxYzkwODUwNWM) + [online ressources](https://github.com/frrrenchies/frrrenchies)
- other languages
- R for DS (slack, tidytuesday)
- ropensci community calls / shinyapp (https://ropensci.shinyapps.io/contributr/)
- engage on twitter ([R for the rest of us](https://twitter.com/rfortherest) + [weareRLadies](https://twitter.com/WeAreRLadies/status/1154698236583698432) + [Mara Averick](https://twitter.com/dataandme) + [maelle Salmon](https://twitter.com/ma_salmon))

Make an example: 

- RUG in GENEVA, R lunches
- R Ladies remote (journal club)






<!------------------------------------------------------->
<!---------------------- ANNEX -------------------------->
<!------------------------------------------------------->
---
class: left, middle, inverse


# Links

---
class: left, middle

# More info

- https://learningstatisticswithr.com/
- https://r4ds.had.co.nz/

---
class: left, middle

# Inspiration

- https://twitter.com/phillynerd/status/1130633179134607360?s=09
- [Ten Quick Tips for Delivering a Programming Lesson](http://third-bit.com/2019/06/15/10-quick-tips-for-delivering-a-programming-lesson.html) by Greg Wilson
- [Curriculum Development](https://twitter.com/dataandme/status/1140390752670965760) by Mine Cetinkaya-Rundel
- Clearify for whom this course is: https://twitter.com/RochelleTerman/status/1126642900006252544
- https://github.com/rstudio/parsons
- https://carpentries.org/blog/2019/06/reproducible-science-r/
- illustrations from http://jules32.github.io/
- [ggplot_flipbook](https://github.com/EvaMaeRey/ggplot_flipbook)
- [demoR](https://github.com/kbodwin/demoR)
- reproducible research: https://annakrystalli.me/talks/r-in-repro-research.html#15
