---
title: "Create a package"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Preparation

1. Create a [github account](https://happygitwithr.com/rstudio-git-github.html)
1. Connect [RStudio to git and github](https://happygitwithr.com/rstudio-git-github.html)
1. Install the `usethis` and the `devtools` package:
  - `install.packages("devtools")`
  - `install.packages("usethis")`

## Create the package skeleton

Run the following command in a R console:

`usethis::create_package("~/tmp/EPFLcorpid")`

This will open a new RStudio window. Head over there. 

Then make sure the package repository will use git by typing: 

```
usethis::use_git()
```

Add a license:

```
usethis::use_mit_license()
```

Run a first check:

```
devtools::check()
```

Then commit all files with git (usually the top-right corner):


Set up git + github:
```
usethis::use_github()
```
This will also update the `DESCRIPTION` file.

Install the package:
```
devtools::install()
```

Add a rmarkdown `README` file that contains some information about how to install your package:

```
usethis::use_readme_rmd()
```

Knit the package, commit it and push all files. 

---
## Add content

### R Code
Add an R-file with the following command:
```
usethis::use_r("epfl_colors.R")
```

For now we will add a function that looks like [this one](https://github.com/sinarueeger/EPFLcorpid_template/blob/master/R/epfl_colors.R).

```{r eval = FALSE}

epfl_colors <- function(name = NULL, message = FALSE)
{

  #color_mat <- tibble::tribble(
  #  ~epfl_color_name, ~human_color_name, ~hex,
  #  "groseille", "dark-red", "#B51F1F" ,
  #  "rouge", "red", "#FF0000"    ,
  #  "canard", "dark-blue", "#007480" ,
  #  "leman", "blue", "#00A79F"   ,
  #  "ardoise", "dark-gray", "#413D3A"  ,
  #  "perle", "gray", "#CAC7C7"
  #)

  color_mat <- structure(
    list(
      epfl_color_name = c("groseille", "rouge", "canard", "leman", "ardoise", "perle"),
      human_color_name = c("dark-red", "red", "dark-blue", "blue", "dark-gray", "gray"),
      hex = c("#B51F1F", "#FF0000", "#007480", "#00A79F", "#413D3A", "#CAC7C7")),
    row.names = c(NA, -6L), class = c("tbl_df", "tbl", "data.frame"))

  ## repeat lines that are blue for greed
  color_mat_green <- color_mat[grep("blue", color_mat$human_color_name), ]
  color_mat_green$human_color_name <- gsub("blue", "green", color_mat_green$human_color_name)

  color_mat <- rbind(color_mat, color_mat_green)

  color_mat_long <- stats::reshape(as.data.frame(color_mat), varying = list(1:2), direction = "long")[, c("epfl_color_name", "hex")]


  if (message) message("EPFL corp id:\n\nUsing colors appropriately is one of the easiest ways to ensure that our documents reflect a coherent visual image or story from EPFL.\n\nAlways balance colors with generous white spaces.
")

  if(is.null(name))
  {
    return(color_mat)
  } else {

    if(any(color_mat_long$epfl_color_name %in% name))
    {
      out <- sapply(name, function(x) color_mat_long[color_mat_long$epfl_color_name == x, "hex"])
    }else {
      stop("The color you choose is not an EPFL color.")
    }

    return(out)

  }


}


```

### Add documentation

```{r eval = FALSE}
#' Get hex color ids for EPFL corporate colors
#'
#' @param name vector containing the color names (see below). If no name is supplied, then the whole color dataframe with two
#' color names and the hex code is returned.
#' @param message logical, displays the note from EPFL corporate identity.
#'
#' @details Color names are: \code{"dark-red"}, \code{"red"},
#' \code{"dark-blue"}, \code{"blue"},  \code{"dark-green"}, \code{"green"},
#' \code{"dark-gray"}, \code{"gray"}, or the the official names \code{"groseille"},
#' \code{"rouge"}, \code{"canard"}, \code{"leman"}, \code{"ardoise"},
#' \code{"perle"}.
#'
#' @references Source: \href{https://www.epfl.ch/campus/services/communication/wp-content/uploads/2019/03/EPFL-brand-guidelines.pdf}{EPFL Brand Guidelines}
#'
#' @return Named character vector or dataframe.
#' @export
#'
#' @examples
#' epfl_colors()
#' epfl_colors("blue")
#' epfl_colors("green")
#' epfl_colors("green", message = TRUE)
#'

```
Run the documentation with `Ctrl + Shift + D`

### Update description

1. `usethis::use_package()`

### Rinse and repeat

The steps are always: 

- Load `Ctrl + Shift + L` (`devtools::load()`)
- Build `Ctrl + Shift + B` (`devtools::install()`)
- Document `Ctrl + Shift + D` (`devtools::document()`)
- Check `Ctrl + Shift + E` (`devtools::check()`)

### Pimp the README file

1. Add a purpose + example
1. Render the file


### Push to your github repository


## Extra stuff


### Add unit tests

1. `usethis::use_testthat()`
1. `usethis::use_test("epfl_colors")`

### Continuous integration

`usethis::use_travis()`

### Add vignette

`usethis::use_vignette()`

### Create a webpage

`pkgdown::build_site()`

### Submitting to CRAN

Read [here](http://r-pkgs.had.co.nz/release.html)


--- 
## Clean up if this was only a demo

```{r, eval = FALSE}
install.packages("fs")
fs::dir_delete("~/tmp/mypackage")
```