---
title: "Meteorite report"
author: <Your Name>
date: "2019-09-07 (updated: `r Sys.Date()`)" 
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    number_sections: true
    highlight: tango

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(dplyr)
# library(readr)

library(maps) ## to visualise spatial data

library(writexl) ## to write out xlsx

```


This report is about **meteorites**, where they fell and when they fell. Data comes from the Meteoritical Society by way of NASA and was contributed to [TidyTuesday](https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-06-11) (week 24) by [Malin Axelsson](https://twitter.com/malinfax).

If you want to read more about the classification of meteorites, see [this Wikipedia](https://en.wikipedia.org/wiki/Meteorite_classification) entry. 


# Import data

```{r import-data}
dat_raw <- readr::read_csv("../data/meteorites.csv")
```

The `meteorites` dataset contains `r nrow(dat_raw)` observations and `r ncol(dat_raw)` variables. The csv file was downloaded from `https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-06-11/meteorites.csv` and stored as `meteorites.csv`. 



# Tidy data

There is no data tidying needed (see [R4DS tidy data](https://r4ds.had.co.nz/tidy-data.html)).


# Transform data

```{r transform-data}
dat <- dat_raw %>% 
## 1) Remove rows that have year as NA
  filter(!is.na(year)) %>%
## 2) remove name_type Relict cases
  filter(name_type != "Relict") %>%
## 3) remove mass min
  filter(mass > 0) %>%
## 4) remove years with more than 2019
  filter(year <= 2019) %>%
## 5) create a variable century
  mutate(century = droplevels(cut(year, seq(from = 0, to = 2019, by = 100), right = FALSE, dig.lab = 4))) %>% 
## 6) create a variable decade
  mutate(decade = droplevels(cut(year, seq(from = 0, to = 2019, by = 10), right = FALSE, dig.lab = 4)))

```

# Descriptive analysis

```{r desc-analysis}
skimr::skim(dat) %>% 
  skimr::kable(format = 'html') 

```


## Histogram

```{r dataviz-hist}
ggplot(data = dat) + geom_histogram(aes(year)) + xlim(1800, NA)
```

## Scatterplot

```{r dataviz-point}
ggplot(data = dat) + geom_line(aes(year, mass)) + xlim(1800, NA)
```


## Map

```{r dataviz-map-like, fig.height=7, fig.width=7}
ggplot(data = dat) + geom_point(aes(long, lat))
ggplot(data = dat) + geom_point(aes(long, lat)) + facet_wrap(~ century)
```



```{r dataviz-map, fig.height=7, fig.width=7, message=FALSE, warning=FALSE}

world_map <- map_data("world")
ggplot() +
  geom_polygon(data = world_map, aes(x = long, y = lat, group = group), fill="lightgray", colour = "white") + 
  geom_point(data = dat, aes(long, lat), color="red", alpha=0.5) + 
  theme_void()

```


<!------------
## Animation

```{r dataviz-animation, eval=FALSE, fig.height=7, fig.width=7, include=FALSE}
library(gganimate) ## to animate the map

world_map <- map_data("world")
ggplot(data = dat) +
  geom_polygon(data = world_map, aes(x = long, y = lat, group = group), fill="lightgray", colour = "white") + 
  geom_point(data = dat, aes(long, lat), color="red", alpha=0.5) +
  transition_states(decade, transition_length = 40, state_length = 40) + 
  labs(title = "Decade: {closest_state}") +
  NULL

```
---------->

# Export data

Store transformed data

```{r write-data}

## as text file
readr::write_delim(dat, path = "../out/meteorites_transformed.csv", delim = " " )

## as Excel file
?write_xlsx
write_xlsx(dat, path = "../out/meteorites_transformed.xlsx")

```




# More infos about RMarkdown

If you want to learn more about RMarkdown, see:

- [this cheatsheet](https://resources.rstudio.com/rstudio-cheatsheets/rmarkdown-2-0-cheat-sheet).
- https://holtzy.github.io/Pimp-my-rmd/
- https://rmd4sci.njtierney.com/

